---
title: "Celine-flexdashboard"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 5
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(rvest)
library(httr)
library(ggplot2)
library(readr)
library(plotly)
library(flexdashboard)
library(ggthemes)
library(shinyWidgets)
library(reshape2)
library(gt)




# Packages required for maps 
library(tidyverse)
library(mdsr)      # package associated with our MDSR book
library(poliscidata)
library(htmltools)
library(glue)
library(maps)
library(viridis)
library(statebins)
library(leaflet)
library(dplyr)
```

```{r, echo = FALSE}
artists <- read_csv("global_music.csv")

```

# About

## Column {data-width = 400}

#### Background Information

Our data was acquired though the [MusicBrainz API](https://musicbrainz.org/doc/MusicBrainz_API) and the files for this can be found in our [GitHub](https://github.com/CelineeUWC/SDS_MP2).

# Maps

## Column {data-width = 400}

*Lets take a look at some maps*

```{r, echo = FALSE}

#here i got the average coordinate for each country 
country_coords <- map_data("world") |>
  
  group_by(region) |>
  
  summarize(
    Longitude = mean(long, na.rm = TRUE),
    Latitude = mean(lat, na.rm = TRUE)
  )

head(country_coords)


#here i cleaned the country names 

```

```{r}

artists <- artists |>
  mutate(
    Country = recode(
      Country,
      "united states" = "usa",
      "united kingdom" = "uk",
      "england" = "uk",
      "los angeles" = "usa",
      
    )
  )

artists1
```

```{r}
# more name cleaning 
artists$Country <- tolower(artists$Country)
country_coords$region <- tolower(country_coords$region)
```

```{r}

artists_map <- artists %>%
  
  left_join(
    country_coords,
    by = c("Country" = "region")
  )
```

## Column

currently my artits data does not have the longigute and latitude i need to get it form some where else

```{r}
ggplot(artists_map, aes(Longitude, Latitude))+
  geom_polygon(color = "white") +
  coord_fixed(1.3) +
  scale_fill_viridis_c(na.value = "gray90") +
  labs(
    title = "Global Music Artists by Country",
    fill = "Artists"
  ) +
  theme_minimal()
```

## Column

Here we will attempt to join map data with the global music artists data that we acquired through the MusicBrainz API database.

# Artist Analysis

Column {.sidebar data-width=260}
--------------------------------------------------------------------------------
### Controls

```{r,echo=FALSE}
inputPanel(
selectInput(
  "n_breaks",
  label = "Country:",
  choices = c("lebanon","mexico","usa","uk","canada")
)
)
```

Column {data-width=200}
--------------------------------------------------------------------------------

### Map

```{r,echo=FALSE}

output$map <- renderLeaflet({

  if (input$n_breaks == "lebanon") {

    leaflet() |>
      addTiles() |>
      setView(35.969155, 33.8413248, zoom = 6) |>
      addPopups(35.969155, 33.8413248, "Welcome to Lebanon")

  } else if (input$n_breaks == "mexico") {

    leaflet() |>
      addTiles() |>
      setView(-102.5528, 23.6345, zoom = 5) |>
      addPopups(-102.5528, 23.6345, "Welcome to Mexico")

  }else if (input$n_breaks == "usa") {

    leaflet() |>
      addTiles() |>
      setView(-121.6253101,48.7433334, zoom = 5) |>
      addPopups(-121.6253101,48.7433334, "Welcome to United States")
  }else if (input$n_breaks == "uk") {

    leaflet() |>
      addTiles() |>
      setView(-4.0987501,55.5581317, zoom = 5) |>
      addPopups(-4.0987501,55.5581317, "Welcome to United Kindgom")
  } else if (input$n_breaks == "canada") {

    leaflet() |>
      addTiles() |>
      setView(-89.756176, 65.2931221, zoom = 5) |>
      addPopups(-89.756176,65.2931221, "Welcome to Canada")
  } 

})
```

```{r,echo=FALSE}
leafletOutput("map", height = 400)
```

Column {data-width=200}
--------------------------------------------------------------------------------
### USer Rating Plot

```{r,echo=FALSE}
renderPlotly({

  if (input$n_breaks == "lebanon") {
    artists |> 
      filter(Country == "lebanon") |>
     plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
  } else if (input$n_breaks == "mexico") {
    artists |> 
      filter(Country == "mexico") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )

  } else if (input$n_breaks == "usa") {
    artists |> 
      filter(Country == "united states") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
  }  else if (input$n_breaks == "uk") {
    artists |> 
      filter(Country == "united kingdom") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
    
  } else if (input$n_breaks == "canada") {
    artists |> 
      filter(Country == "canada") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
  }

})

```

### Collabration Plot

```{r, echo=FALSE}
renderPlotly({

  if (input$n_breaks == "lebanon") {
    artists |>
      filter(Country == "Lebanon") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  } else if (input$n_breaks == "mexico") {
   artists |> 
      filter(Country == "mexico") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  } else if (input$n_breaks == "usa") {
    artists |> 
      filter(Country == "united states") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  }  else if (input$n_breaks == "uk") {
    artists |> 
      filter(Country == "united kingdom") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  } else if (input$n_breaks == "canada") {
   artists |> 
      filter(Country == "canada") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
}

})

```



# Collabs

# Conclusion
