---
title: "Global Artists"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 5
    navbar:
      - { title: "GitHub", href: "https://github.com/CelineeUWC/SDS_MP2", align: right, icon: "fa-github" }
    source_code: embed
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(stringr)
library(ggplot2)
library(readr)
library(plotly)
library(flexdashboard)
library(shinyWidgets)
library(gt)
library(shiny)
library(maps)
library(viridis)
library(leaflet)
library(htmltools)
library(glue)
library(tidygeocoder) #new package we found for hometown coordinates 
```


```{r load-and-clean, echo=FALSE}
# load data
artists <- read_csv("global_music.csv") |>
  select(-`...1`)   # drop the auto-index column R adds on read

# standardize country names
# We convert everything to lowercase FIRST, then rename, so there are no
# capitalization mismatches when we later join to map_data("world").
artists <- artists |>
  mutate(Country = tolower(trimws(Country))) |>
  mutate(Country = recode(Country,
    "united states"       = "usa",
    "united kingdom"      = "uk",
    "england"             = "uk",
    "scotland"            = "uk",
    "wales"               = "uk",
    "northern ireland"    = "uk",
    "los angeles"         = "usa",
    "new york"            = "usa",
    "south korea"         = "south korea",
    "republic of ireland" = "ireland"
  ))

# country  lookup from the maps package
# map_data("world") gives one row per polygon vertex, so we collapse each
# country down to a single average lat/lon to use as a map marker position.
country_coords <- map_data("world") |>
  mutate(region = tolower(region)) |>
  group_by(region) |>
  summarize(
    Longitude = mean(long, na.rm = TRUE),
    Latitude  = mean(lat,  na.rm = TRUE),
    .groups   = "drop"
  )

# join coordinates onto the artist rows 
# artists whose country didn't match any map region will get NA coords and
# simply won't appear as dots on the leaflet map
artists_map <- artists |>
  left_join(country_coords, by = c("Country" = "region"))

# in load-and-clean chunk, run once, then always read from cache
if (!file.exists("artists_geocoded.csv")) {

  # create better addresses
  artists_map <- artists_map |>
    mutate(
      full_location = paste(Hometown, Country, sep = ", ")
    )

  # geocode
  artists_map <- artists_map |>
    geocode(
      address = full_location,
      method = "arcgis",
      long = Hometown_long,
      lat = Hometown_lat
    )

  write_csv(artists_map, "artists_geocoded.csv")

} else {

  artists_map <- read_csv("artists_geocoded.csv")

}

# pre-build filter choice vectors (used by every sidebar)
all_genres    <- c("All", sort(unique(na.omit(artists$Top_Genre))))
all_types     <- c("All", sort(unique(na.omit(artists$Type))))
all_genders   <- c("All", sort(unique(na.omit(artists$Gender))))
all_countries <- c("All", sort(unique(na.omit(artists$Country))))
age_range     <- range(artists$Age, na.rm = TRUE)
```

Home
================================================================================

Column {data-width=480}
--------------------------------------------------------------------------------

### About This Project

Welcome to the **Global Artists Dashboard** a data-driven exploration of
some of our favorite music artists from around the world.

---

#### Data Source

All artist data was collected from the
[**MusicBrainz API**](https://musicbrainz.org/doc/MusicBrainz_API), an open
music encyclopedia that records metadata about artists, releases, and
recordings worldwide. This dataset was constructed by manually selecting 96 artists to query through the MusicBrainz API. As a result, findings reflect patterns within this curated sample and cannot be generalized to the broader global music landscape. Future work could use a random or stratified sample to reduce selection bias.

---

#### How We Built the Dataset

| Step | Action |
|------|--------|
| 1 | Queried the MusicBrainz REST API across multiple genres and regions |
| 2 | Extracted artist metadata (ratings, tags, type, area) |
| 3 | Derived **Age** from career-begin dates |
| 4 | Merged with `map_data("world")` for geographic coordinates |
| 5 | Cleaned and standardized country names for joining |

---

#### Variables in the Dataset

| Variable | Type | Description |
|----------|------|-------------|
| `Artist` | Categorical | Artist or group name |
| `Type` | Categorical | *Person* (solo) or *Group* |
| `Gender` | Categorical | Gender of solo artist |
| `Top_Genre` | Categorical | Primary musical genre |
| `Country` | Categorical | Country of origin |
| `Hometown` | Categorical | City/region of origin |
| `Is_Active` | Binary Categorical | Whether artist is still active |
| `User_Rating` | Numerical | Avg. user rating (0–5) |
| `Rating_Votes` | Numerical | Number of user votes |
| `Collab_Count` | Numerical | Count of credited collaborations |
| `Age` | Numerical | Years since career began |
| `Album_Count` | Numerical | Number of albums |
| `Recording_Count` | Numerical | Number of recordings |

The full code and raw data live on our
[**GitHub Repository →**](https://github.com/CelineeUWC/SDS_MP2)


Column {data-width=520}
--------------------------------------------------------------------------------


### Dataset at a Glance

```{r home-kpi}
n_artists   <- nrow(artists)
n_countries <- n_distinct(artists$Country,   na.rm = TRUE)
n_genres    <- n_distinct(artists$Top_Genre, na.rm = TRUE)
avg_rating  <- round(mean(artists$User_Rating, na.rm = TRUE), 2)

tibble(
  Metric = c("Total Artists", "Countries Represented",
             "Genres Represented", "Avg. User Rating"),
  Value  = c(
    format(n_artists, big.mark = ","),
    as.character(n_countries),
    as.character(n_genres),
    paste0(avg_rating, " / 5")
  )
) |>
  gt() |>
  tab_header(title = "Dataset Overview") |>
  cols_align("center", columns = Value) 
```

### Sample of the Data

```{r home-table}
artists |>
  select(Artist, Country, Top_Genre, Type, Gender,
         Age, User_Rating, Rating_Votes, Collab_Count, Is_Active) |>
  slice_head(n = 15) |>
  gt() |>
  cols_label(
    Top_Genre    = "Genre",
    User_Rating  = "Rating",
    Rating_Votes = "Votes",
    Collab_Count = "Collabs",
    Is_Active    = "Active"
  ) |>
  fmt_number(columns = User_Rating, decimals = 2)
```


About
================================================================================



### Research Questions

This project investigates **what makes a music artist popular** by analyzing
artist characteristics, geography, age, and collaboration behavior.

---

#### Popularity & Ratings
1. What factors are most associated with higher user ratings?
2. How do genre, type (solo vs group), and country relate to ratings?

####  Age & Career
3. At what age do most highly-rated **solo artists** begin their careers?
4. How are age and genre related in this data set?

####  Geography
5. Does an artist's location affect their popularity?

####  Collaboration
6. Are younger artists collaborating more than older artists?
7. How does collaboration vary by artist type (group vs person)?


### Summary Statistics by Artist Type

```{r about-summary}
artists |>
  filter(!is.na(Type)) |>
  group_by(Type) |>
  summarize(
    Count       = n(),
    Avg_Rating  = round(mean(User_Rating,  na.rm = TRUE), 2),
    Median_Age  = median(Age,              na.rm = TRUE),
    Avg_Collabs = round(mean(Collab_Count, na.rm = TRUE), 1),
    Avg_Albums  = round(mean(Album_Count,  na.rm = TRUE), 1),
    .groups     = "drop"
  ) |>
  gt() |>
  cols_label(
    Avg_Rating  = "Avg Rating",
    Median_Age  = "Median Age",
    Avg_Collabs = "Avg Collabs",
    Avg_Albums  = "Avg Albums"
  )
```


Maps
================================================================================

Column {.sidebar data-width=260}
--------------------------------------------------------------------------------

**Map Filters**

```{r map-sidebar}
selectInput("map_genre", "Genre:",
            choices = all_genres, selected = "All")

selectInput("map_type", "Artist Type:",
            choices = all_types, selected = "All")

selectInput("map_active", "Active Status:",
            choices = c("All", "TRUE", "FALSE"), selected = "All")

sliderInput("map_rating", "Min. User Rating:",
            min = 0, max = 5, value = 0, step = 0.25)

hr()
p(em("Circle size = rating. Click any marker for full details."))
```

Column {data-width=700}
--------------------------------------------------------------------------------

### Interactive World Map of Artists

```{r map-leaflet}
# reactive filter
# every time a sidebar input changes, R re-runs this reactive block and the
# map below automatically redraws with only the matching rows.
map_data_r <- reactive({
  df <- artists_map |> 
    filter(!is.na(Hometown_long), !is.na(Hometown_lat))
  if (input$map_genre  != "All") df <- df |> filter(Top_Genre == input$map_genre)
  if (input$map_type   != "All") df <- df |> filter(Type      == input$map_type)
  if (input$map_active != "All")
    df <- df |> filter(as.character(Is_Active) == input$map_active)
  df <- df |> filter(is.na(User_Rating) | User_Rating >= input$map_rating)
  df
})

renderLeaflet({
  df  <- map_data_r()
  # colorNumeric maps a continuous variable (User_Rating 0–5) to a colour ramp.
  pal <- colorNumeric("YlOrRd", domain = c(0, 5), na.color = "#aaaaaa")

  leaflet(df) |>
    addProviderTiles(providers$CartoDB.Positron) |>  # clean greyscale base map
    setView(lng = 10, lat = 25, zoom = 2) |>
    addCircleMarkers(
      # jitter slightly so artists in the same country don't all stack exactly
      lng         = ~jitter(Hometown_long, factor = 10), 
      lat         = ~jitter(Hometown_lat, factor = 10),
      # larger circle = higher rating
      radius      = ~ifelse(is.na(User_Rating), 5, 3 + User_Rating * 2),
      fillColor   = ~pal(User_Rating),
      fillOpacity = 0.85,
      stroke      = TRUE,
      weight      = 0.8,
      color       = "white",
      # html popup shown when user clicks a marker
      popup = ~paste0(
        "<b style='font-size:14px'>", Artist, "</b><br>",
        "<hr style='margin:4px 0'>",
        " <b>Country:</b> ",  str_to_title(Country), "<br>",
        "️ <b>Hometown:</b> ", ifelse(is.na(Hometown),  "Unknown", Hometown),  "<br>",
        " <b>Genre:</b> ",    ifelse(is.na(Top_Genre), "Unknown", Top_Genre), "<br>",
        " <b>Type:</b> ",     ifelse(is.na(Type),      "Unknown", Type),      "<br>",
        " <b>Rating:</b> ",   ifelse(is.na(User_Rating), "N/A",
                                       paste0(round(User_Rating, 2), " / 5")),  "<br>",
        " <b>Age:</b> ",      ifelse(is.na(Age), "N/A", Age),                 "<br>",
        " <b>Collabs:</b> ",  ifelse(is.na(Collab_Count), "N/A", Collab_Count)
      ),
      label = ~Artist   # tooltip shown on hover (no click needed)
    ) |>
    addLegend("bottomright",
              pal     = pal,
              values  = ~User_Rating,
              title   = "User Rating",
              opacity = 0.85)
})
```

Within our sample, most artists cluster in the USA, UK, and Canada, with smaller concentrations in Lebanon and Mexico. Circle size (rating) doesn't appear to concentrate in any single region, suggesting that within this curated group, geography alone is not a strong predictor of user rating.

Country Analysis
==========================

Column {.sidebar data-width=260}
--------------------------------------------------------------------------------
### Controls

```{r,echo=FALSE}
inputPanel(
selectInput(
  "n_breaks",
  label = "Country:",
  choices = c("lebanon","mexico","usa","uk","canada")
)
)
```



```{r,echo=FALSE}
artist_leb <- artists_map |>
  select(Artist,Country, Hometown_lat, Hometown_long, Hometown)|>
  filter(Country == "lebanon")

artist_mex <-  artists_map |>
  select(Artist,Country, Hometown_lat,Hometown_long, Hometown)|>
  filter(Country == "mexico")

artist_usa <-  artists_map |>
  select(Artist,Country,  Hometown_lat, Hometown_long, Hometown)|>
  filter(Country == "usa") 

artist_uk <-  artists_map |>
  select(Artist,Country,  Hometown_lat, Hometown_long, Hometown)|>
  filter(Country == "uk")
artist_can <-  artists_map |>
  select(Artist,Country,  Hometown_lat, Hometown_long, Hometown)|>
  filter(Country == "canada")

```

Column {data-width=200}
--------------------------------------------------------------------------------

### Map

```{r,echo=FALSE}

output$map <- renderLeaflet({

  if (input$n_breaks == "lebanon") {

    leaflet() |>
      addTiles() |>
      setView(lng = mean(artist_leb$Hometown_long), lat = mean(artist_leb$Hometown_lat), zoom = 6) |>
      addPopups(35.969155, 33.8413248, "Welcome to Lebanon, Lebanese music is a vibrant mix of traditional Levantine folk and modern, high-energy pop, often featuring heavy rhythmic beats from the darbuka (goblet drum) and riq (tambourine). It is a massive regional hub for both Arabic pop and genres like deep house, with Beirut renowned for its thriving nightlife and rooftop club scene") |>
      addCircleMarkers(data = artist_leb,
                      lng = ~Hometown_long,
                      lat = ~Hometown_lat,
                       color = "green",
                      popup = paste0("Name: ", artist_leb$Artist))

  } else if (input$n_breaks == "mexico") {

    leaflet() |>
      addTiles() |>
      setView(lng = mean(artist_mex$Hometown_long), lat = mean(artist_mex$Hometown_lat), zoom = 6) |>
      addPopups(-102.5528, 23.6345, "Welcome to Mexico,Mexican music is a vibrant, diverse, and historical blend of indigenous, Spanish, and African influences, heavily featuring strings, brass, and storytelling. It ranges from 500-year-old folk styles to modern, chart-topping rancheras, often highlighting themes of love, loss, and national pride") |>
      addCircleMarkers(data = artist_mex,
                      lng = ~Hometown_long,
                      lat = ~Hometown_lat,
                       color = "pink",
                      popup = paste0("Name: ", artist_mex$Artist))


  }else if (input$n_breaks == "usa") {

    leaflet() |>
      addTiles() |>
       setView(lng = mean(artist_usa$Hometown_long), lat = mean(artist_usa$Hometown_lat), zoom = 6) |>
      addPopups(-121.6253101,48.7433334, "Welcome to United States, American music history is full of surprising trivia, such as the fact that iconic bands like Led Zeppelin and R.E.M. never achieved a #1 single in the US, while Rihanna has ten. The best-selling instrument in the US is the harmonica, and the cast of the show Glee has charted more songs than The Beatles")|>
      addCircleMarkers(data = artist_usa,
                      lng = ~Hometown_long,
                      lat = ~Hometown_lat,
                       color = "blue",
                      popup = paste0("Name: ", artist_usa$Artist))

  }else if (input$n_breaks == "uk") {

    leaflet() |>
      addTiles() |>
      setView(lng = mean(artist_uk$Hometown_long), lat = mean(artist_uk$Hometown_lat), zoom = 6) |>
      addPopups(-4.0987501,55.5581317, "Welcome to United Kindgom, The UK music scene has a rich, compact history that has shaped global pop culture, featuring pioneering punk scenes, extensive folk traditions, and enduring chart rivalries. Elton John’s Candle in the Wind '97 is the best-selling single in UK history, while Lewis Capaldi holds the top-streamed song record") |>
      addCircleMarkers(data = artist_uk,
                      lng = ~Hometown_long,
                      lat = ~Hometown_lat,
                       color = "orange",
                      popup = paste0("Name: ", artist_uk$Artist))

  } else if (input$n_breaks == "canada") {

    leaflet() |>
      addTiles() |>
       setView(lng = mean(artist_can$Hometown_long), lat = mean(artist_can$Hometown_lat), zoom = 6) |>
      addPopups(-89.756176,65.2931221, "Welcome to Canada,Canada is a massive exporter of musical talent, often ranking as the world's third-largest behind the US and UK. Notable quirks include Bryan Adams photographing the Queen, 1970s punk band D.O.A. coining the term hardcore and Toronto pianist Ruth Lowe writing Frank Sinatra's breakout hit I'll Never Smile Again") |>
      addCircleMarkers(data = artist_can,
                      lng = ~Hometown_long,
                      lat = ~Hometown_lat,
                       color = "black",
                      popup = paste0("Name: ", artist_can$Artist))

  } 

})
```

```{r,echo=FALSE}
leafletOutput("map", height = 400)
```

Column {data-width=200}
--------------------------------------------------------------------------------
### User Rating Plot

```{r,echo=FALSE}
renderPlotly({

  if (input$n_breaks == "lebanon") {
    artists |> 
      filter(Country == "lebanon") |>
     plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
  } else if (input$n_breaks == "mexico") {
    artists |> 
      filter(Country == "mexico") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )

  } else if (input$n_breaks == "usa") {
    artists |> 
      filter(Country == "usa") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
  }  else if (input$n_breaks == "uk") {
    artists |> 
      filter(Country == "uk") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
    
  } else if (input$n_breaks == "canada") {
    artists |> 
      filter(Country == "canada") |>
      plot_ly(
        x = ~Artist,
        y = ~User_Rating,
        type = "bar",
        color = ~Gender,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Artist"),
        yaxis = list(title = "User Rating"),
        barmode = "group"
      )
  }

})

```

### Collabration Plot

```{r, echo=FALSE}
renderPlotly({

  if (input$n_breaks == "lebanon") {
    artists |>
      filter(Country == "lebanon") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  } else if (input$n_breaks == "mexico") {
   artists |> 
      filter(Country == "mexico") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  } else if (input$n_breaks == "usa") {
    artists |> 
      filter(Country == "usa") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  }  else if (input$n_breaks == "uk") {
    artists |> 
      filter(Country == "uk") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
  } else if (input$n_breaks == "canada") {
   artists |> 
      filter(Country == "canada") |>
      plot_ly(
        x = ~Age,
        y = ~Collab_Count,
        type = "bar",
        color = ~Artist,
        text = ~Artist,
        hovertemplate = "<b>%{text}</b><br>Rating: %{y}<extra></extra>"
      ) |>
      layout(
        xaxis = list(title = "Age"),
        yaxis = list(title = "Collab Count"),
        barmode = "group"
      )
}

})

```

Among the artists in our sample, rating patterns differ by country. US and UK artists show the widest spread of ratings, reflecting the diversity of genres represented in those selections. Lebanese artists in our sample tend to cluster in a narrower rating band. Gender distribution in ratings also varies by country — in some countries the higher-rated artists skew male, while in others ratings are more evenly spread across genders.

Collabs
==========================


Column {.sidebar data-width=260}
--------------------------------------------------------------------------------

**Collaboration Filters**

```{r collab-sidebar}
sliderInput("col_age", "Artist Age Range:",
            min   = floor(age_range[1]),
            max   = ceiling(age_range[2]),
            value = c(floor(age_range[1]), ceiling(age_range[2])),
            step  = 1)

selectInput("col_type", "Artist Type:",
            choices = all_types, selected = "All")

selectInput("col_genre", "Genre:",
            choices = all_genres, selected = "All")

selectInput("col_active", "Active Status:",
            choices = c("All", "TRUE", "FALSE"), selected = "All")
```

Column {data-width=580}
--------------------------------------------------------------------------------

###  Age vs. Collaboration Count

```{r collab-scatter}
col_data <- reactive({
  df <- artists |>
    filter(!is.na(Age), !is.na(Collab_Count)) |>
    filter(Age >= input$col_age[1], Age <= input$col_age[2])
  if (input$col_type   != "All") df <- df |> filter(Type      == input$col_type)
  if (input$col_genre  != "All") df <- df |> filter(Top_Genre == input$col_genre)
  if (input$col_active != "All")
    df <- df |> filter(as.character(Is_Active) == input$col_active)
  df
})

# scatterplot with an optional linear trend line.
# The trend line is built with lm() and added via add_lines().
# We guard with nrow >= 5 so the model doesn't crash on tiny subsets.
renderPlotly({
  df <- col_data()

  p <- plot_ly(df,
               x     = ~Age,
               y     = ~Collab_Count,
               type  = "scatter",
               mode  = "markers",
               color = ~Type,
               text  = ~Artist,
               hovertemplate = "<b>%{text}</b><br>Age: %{x}<br>Collabs: %{y}<extra></extra>",
               marker = list(size = 8, opacity = 0.7))

  if (nrow(df) >= 5 && sd(df$Age, na.rm = TRUE) > 0) {
    fit <- lm(Collab_Count ~ Age, data = df)
    p <- p |>
      add_lines(x          = ~Age,
                y          = fitted(fit),
                name       = "Linear trend",
                line       = list(color = "red", dash = "dash", width = 2),
                inherit    = FALSE,
                showlegend = TRUE)
  }

  p |> layout(
    xaxis  = list(title = "Artist Age (years)"),
    yaxis  = list(title = "Collaboration Count"),
    legend = list(title = list(text = "Type"))
  )
})
```

###  Collab Count vs. User Rating

```{r collab-rating-scatter}
# do artists who collaborate more also get rated higher?
# This directly addresses research question 6.
renderPlotly({
  col_data() |>
    filter(!is.na(User_Rating)) |>
    plot_ly(x     = ~Collab_Count,
            y     = ~User_Rating,
            type  = "scatter",
            mode  = "markers",
            color = ~Type,
            text  = ~Artist,
            hovertemplate = "<b>%{text}</b><br>Collabs: %{x}<br>Rating: %{y:.2f}<extra></extra>",
            marker = list(size = 7, opacity = 0.65)) |>
    layout(
      xaxis  = list(title = "Collaboration Count"),
      yaxis  = list(title = "User Rating (0–5)"),
      legend = list(title = list(text = "Type"))
    )
})
```

Column {data-width=420}
--------------------------------------------------------------------------------

###  Collab Count by Artist Type

```{r collab-box}
# boxplot comparing collab distributions for Person vs Group.
# "outliers" shows individual dots for extreme values.
renderPlotly({
  col_data() |>
    filter(!is.na(Type)) |>
    plot_ly(y         = ~Collab_Count,
            color     = ~Type,
            type      = "box",
            colors    = c("blue", "hotpink"),
            boxpoints = "outliers") |>
    layout(
      yaxis  = list(title = "Collaboration Count"),
      xaxis  = list(title = "Artist Type"),
      legend = list(title = list(text = "Type"))
    )
})
```

Among the artists in our sample, the age vs. collaboration scatter shows a slight positive trend — older artists tend to have higher collaboration counts, likely reflecting more time in the industry. However, the trend line slope is shallow, meaning age alone explains only a small portion of the variation.


The collab count vs. user rating scatter shows no strong relationship within our sample, suggesting that collaborating more does not reliably translate to higher ratings for either groups or solo artists.


The boxplot most directly addresses research question 7: among our sampled artists, groups had a median collab count of 17 vs. 4 for solo artists, a notable difference that may reflect how group credits are recorded in MusicBrainz.

Conclusion
==========================

### Conclusion 
Taken together, these findings suggest that within our sample of 96 manually selected artists, artist type (group vs. solo) is the strongest differentiator in collaboration behavior, while geography and age play more modest roles. Because artists were hand-picked rather than randomly sampled, these patterns should not be generalized to global music trends, they reflect characteristics of this particular curated group. Future work using a randomized or stratified sample would allow for stronger conclusions.

### References
- [MusicBrainz API](https://musicbrainz.org/doc/MusicBrainz_API)
- [flexdashboard](https://pkgs.rstudio.com/flexdashboard/)
- [Leaflet for R](https://rstudio.github.io/leaflet/)
- [plotly for R](https://plotly.com/r/)
- [GitHub Repository](https://github.com/CelineeUWC/SDS_MP2)