---
title: "Global Artists"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 5
    navbar:
      - { title: "GitHub", href: "https://github.com/CelineeUWC/SDS_MP2", align: right, icon: "fa-github" }
    source_code: embed
---



```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(stringr)
library(ggplot2)
library(readr)
library(plotly)
library(flexdashboard)
library(shinyWidgets)
library(gt)
library(shiny)
library(maps)
library(viridis)
library(leaflet)
library(htmltools)
library(glue)
```


```{r load-and-clean, echo=FALSE}
# ── Load data ──────────────────────────────────────────────────────────────────
artists <- read_csv("global_music.csv") |>
  select(-`...1`)   # drop the auto-index column R adds on read

# standardise country names
# We convert everything to lowercase FIRST, then recode, so there are no
# capitalisation mismatches when we later join to map_data("world").
artists <- artists |>
  mutate(Country = tolower(trimws(Country))) |>
  mutate(Country = recode(Country,
    "united states"       = "usa",
    "united kingdom"      = "uk",
    "england"             = "uk",
    "scotland"            = "uk",
    "wales"               = "uk",
    "northern ireland"    = "uk",
    "los angeles"         = "usa",
    "new york"            = "usa",
    "south korea"         = "south korea",
    "republic of ireland" = "ireland"
  ))

# country  lookup from the maps package
# map_data("world") gives one row per polygon vertex, so we collapse each
# country down to a single average lat/lon to use as a map marker position.
country_coords <- map_data("world") |>
  mutate(region = tolower(region)) |>
  group_by(region) |>
  summarize(
    Longitude = mean(long, na.rm = TRUE),
    Latitude  = mean(lat,  na.rm = TRUE),
    .groups   = "drop"
  )

# join coordinates onto the artist rows 
# artists whose country didn't match any map region will get NA coords and
# simply won't appear as dots on the leaflet map
artists_map <- artists |>
  left_join(country_coords, by = c("Country" = "region"))

# pre-build filter choice vectors (used by every sidebar)
all_genres    <- c("All", sort(unique(na.omit(artists$Top_Genre))))
all_types     <- c("All", sort(unique(na.omit(artists$Type))))
all_genders   <- c("All", sort(unique(na.omit(artists$Gender))))
all_countries <- c("All", sort(unique(na.omit(artists$Country))))
age_range     <- range(artists$Age, na.rm = TRUE)
```

Home
================================================================================

Column {data-width=480}
--------------------------------------------------------------------------------

### About This Project

Welcome to the **Global Artists Dashboard** a data-driven exploration of
some of our favorite music artists from around the world.

---

#### Data Source

All artist data was collected from the
[**MusicBrainz API**](https://musicbrainz.org/doc/MusicBrainz_API), an open
music encyclopedia that records metadata about artists, releases, and
recordings worldwide.

---

#### How We Built the Dataset

| Step | Action |
|------|--------|
| 1 | Queried the MusicBrainz REST API across multiple genres and regions |
| 2 | Extracted artist metadata (ratings, tags, type, area) |
| 3 | Derived **Collab_Count** by counting credited artist relationships |
| 4 | Derived **Age** from career-begin dates |
| 5 | Merged with `map_data("world")` centroids for geographic coordinates |
| 6 | Cleaned and standardised country names for joining |

---

#### Variables in the Dataset

| Variable | Type | Description |
|----------|------|-------------|
| `Artist` | Categorical | Artist or group name |
| `Type` | Categorical | *Person* (solo) or *Group* |
| `Gender` | Categorical | Gender of solo artist |
| `Top_Genre` | Categorical | Primary musical genre |
| `Country` | Categorical | Country of origin |
| `Hometown` | Categorical | City/region of origin |
| `Is_Active` | Binary Categorical | Whether artist is still active |
| `User_Rating` | Numerical | Avg. user rating (0–5) |
| `Rating_Votes` | Numerical | Number of user votes |
| `Collab_Count` | Numerical | Count of credited collaborations |
| `Age` | Numerical | Years since career began |
| `Album_Count` | Numerical | Number of albums |
| `Recording_Count` | Numerical | Number of recordings |

The full code and raw data live on our
[**GitHub Repository →**](https://github.com/CelineeUWC/SDS_MP2)

About
================================================================================


Column {data-width=420}
--------------------------------------------------------------------------------

### Research Questions

This project investigates **what makes a music artist popular** by analysing
artist characteristics, geography, age, and collaboration behaviour.

---

#### Popularity & Ratings
1. What factors are most associated with higher user ratings?
2. How do genre, type (solo vs group), and country relate to ratings?

####  Age & Career
3. At what age do most highly-rated **solo artists** begin their careers?
4. How are age and genre related in this dataset?

####  Geography
5. Does an artist's location affect their popularity?

####  Collaboration
6. Are younger artists collaborating more than older artists?
7. How does collaboration vary by artist type (group vs person)?

Column {data-width=580}
--------------------------------------------------------------------------------

### Summary Statistics by Artist Type

```{r about-summary}
artists |>
  filter(!is.na(Type)) |>
  group_by(Type) |>
  summarize(
    Count       = n(),
    Avg_Rating  = round(mean(User_Rating,  na.rm = TRUE), 2),
    Median_Age  = median(Age,              na.rm = TRUE),
    Avg_Collabs = round(mean(Collab_Count, na.rm = TRUE), 1),
    Avg_Albums  = round(mean(Album_Count,  na.rm = TRUE), 1),
    .groups     = "drop"
  ) |>
  gt() |>
  cols_label(
    Avg_Rating  = "Avg Rating",
    Median_Age  = "Median Age",
    Avg_Collabs = "Avg Collabs",
    Avg_Albums  = "Avg Albums"
  )
```


Maps
================================================================================

Column {.sidebar data-width=260}
--------------------------------------------------------------------------------

**Map Filters**

```{r map-sidebar}
selectInput("map_genre", "Genre:",
            choices = all_genres, selected = "All")

selectInput("map_type", "Artist Type:",
            choices = all_types, selected = "All")

selectInput("map_active", "Active Status:",
            choices = c("All", "TRUE", "FALSE"), selected = "All")

sliderInput("map_rating", "Min. User Rating:",
            min = 0, max = 5, value = 0, step = 0.25)

hr()
p(em("Circle size = rating. Click any marker for full details."))
```

Column {data-width=700}
--------------------------------------------------------------------------------

### Interactive World Map of Artists

```{r map-leaflet}
# reactive filter
# every time a sidebar input changes, R re-runs this reactive block and the
# map below automatically redraws with only the matching rows.
map_data_r <- reactive({
  df <- artists_map |> filter(!is.na(Longitude), !is.na(Latitude))
  if (input$map_genre  != "All") df <- df |> filter(Top_Genre == input$map_genre)
  if (input$map_type   != "All") df <- df |> filter(Type      == input$map_type)
  if (input$map_active != "All")
    df <- df |> filter(as.character(Is_Active) == input$map_active)
  df <- df |> filter(is.na(User_Rating) | User_Rating >= input$map_rating)
  df
})

renderLeaflet({
  df  <- map_data_r()
  # colorNumeric maps a continuous variable (User_Rating 0–5) to a colour ramp.
  pal <- colorNumeric("YlOrRd", domain = c(0, 5), na.color = "#aaaaaa")

  leaflet(df) |>
    addProviderTiles(providers$CartoDB.Positron) |>  # clean greyscale base map
    setView(lng = 10, lat = 25, zoom = 2) |>
    addCircleMarkers(
      # jitter slightly so artists in the same country don't all stack exactly
      lng         = ~jitter(Longitude, factor = 0.8),
      lat         = ~jitter(Latitude,  factor = 0.8),
      # larger circle = higher rating
      radius      = ~ifelse(is.na(User_Rating), 5, 3 + User_Rating * 2),
      fillColor   = ~pal(User_Rating),
      fillOpacity = 0.85,
      stroke      = TRUE,
      weight      = 0.8,
      color       = "white",
      # html popup shown when user clicks a marker
      popup = ~paste0(
        "<b style='font-size:14px'>", Artist, "</b><br>",
        "<hr style='margin:4px 0'>",
        " <b>Country:</b> ",  str_to_title(Country), "<br>",
        "️ <b>Hometown:</b> ", ifelse(is.na(Hometown),  "Unknown", Hometown),  "<br>",
        " <b>Genre:</b> ",    ifelse(is.na(Top_Genre), "Unknown", Top_Genre), "<br>",
        " <b>Type:</b> ",     ifelse(is.na(Type),      "Unknown", Type),      "<br>",
        " <b>Rating:</b> ",   ifelse(is.na(User_Rating), "N/A",
                                       paste0(round(User_Rating, 2), " / 5")),  "<br>",
        " <b>Age:</b> ",      ifelse(is.na(Age), "N/A", Age),                 "<br>",
        " <b>Collabs:</b> ",  ifelse(is.na(Collab_Count), "N/A", Collab_Count)
      ),
      label = ~Artist   # tooltip shown on hover (no click needed)
    ) |>
    addLegend("bottomright",
              pal     = pal,
              values  = ~User_Rating,
              title   = "User Rating",
              opacity = 0.85)
})
```



Artist Analysis
==========================

Column {data-width=200}
-------------------------------------

### Controls

```{r,echo=FALSE}
selectInput(
  "n_breaks",
  label = "Country:",
  choices = c("Lebanon","Mexico","US","UK","Canada"),
  selected = "Lebanon"
)
```

Column
-------------------------------------

### Map

```{r,echo=FALSE}
renderLeaflet({

  if(input$n_breaks == "Lebanon"){

    leaflet() |>
      addTiles() |>
      setView(35.969155, 33.8413248, zoom = 13) |>
      addPopups(35.969155, 33.8413248, "Welcome to Lebanon")

  } else if(input$n_breaks == "Mexico"){

    leaflet() |>
      addTiles() |>
      setView(-103.5980383, 24.2114149, zoom = 13) |>
      addPopups(-103.5980383, 24.2114149, "Welcome to Mexico")
  }

})
```

### Plot
```{r,echo=FALSE}
artist_leb <- artists |> filter(Country == "Lebanon")|>
  ggplot(aes(x = Artist, y = User_Rating, fill = Gender)) +
      geom_col()
artist_leb 
artist_mex <- artists |> filter(Country == "Mexico") |>
  ggplot(aes(x = Artist, y = User_Rating, fill = Gender)) +
      geom_col()
artist_mex 
renderPlot({

  if (input$n_breaks == "Lebanon") {
artist_leb 
  } else if (input$n_breaks == "Mexico") {
artist_mex 
  }

})
```


Collabs
==========================


Conclusion
==========================





