---
title: "Global Artists"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 5
    navbar:
      - { title: "GitHub", href: "https://github.com/CelineeUWC/SDS_MP2", align: right, icon: "fa-github" }
    source_code: embed
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(stringr)
library(ggplot2)
library(readr)
library(plotly)
library(flexdashboard)
library(shinyWidgets)
library(gt)
library(shiny)
library(maps)
library(viridis)
library(leaflet)
library(htmltools)
library(glue)
```

```{r load-and-clean, echo=FALSE}
# ── Load data ──────────────────────────────────────────────────────────────────
artists <- read_csv("global_music.csv") |>
  select(-`...1`)   # drop the auto-index column R adds on read

# standardise country names
# We convert everything to lowercase FIRST, then recode, so there are no
# capitalisation mismatches when we later join to map_data("world").
artists <- artists |>
  mutate(Country = tolower(trimws(Country))) |>
  mutate(Country = recode(Country,
    "united states"       = "usa",
    "united kingdom"      = "uk",
    "england"             = "uk",
    "scotland"            = "uk",
    "wales"               = "uk",
    "northern ireland"    = "uk",
    "los angeles"         = "usa",
    "new york"            = "usa",
    "south korea"         = "south korea",
    "republic of ireland" = "ireland"
  ))

# country  lookup from the maps package
# map_data("world") gives one row per polygon vertex, so we collapse each
# country down to a single average lat/lon to use as a map marker position.
country_coords <- map_data("world") |>
  mutate(region = tolower(region)) |>
  group_by(region) |>
  summarize(
    Longitude = mean(long, na.rm = TRUE),
    Latitude  = mean(lat,  na.rm = TRUE),
    .groups   = "drop"
  )

# join coordinates onto the artist rows 
# artists whose country didn't match any map region will get NA coords and
# simply won't appear as dots on the leaflet map
artists_map <- artists |>
  left_join(country_coords, by = c("Country" = "region"))

# pre-build filter choice vectors (used by every sidebar)
all_genres    <- c("All", sort(unique(na.omit(artists$Top_Genre))))
all_types     <- c("All", sort(unique(na.omit(artists$Type))))
all_genders   <- c("All", sort(unique(na.omit(artists$Gender))))
all_countries <- c("All", sort(unique(na.omit(artists$Country))))
age_range     <- range(artists$Age, na.rm = TRUE)
```


Home
================================================================================

Column {data-width=480}
--------------------------------------------------------------------------------

### About This Project

Welcome to the **Global Artists Dashboard** a data-driven exploration of
some of our favorite music artists from around the world.

---

#### Data Source

All artist data was collected from the
[**MusicBrainz API**](https://musicbrainz.org/doc/MusicBrainz_API), an open
music encyclopedia that records metadata about artists, releases, and
recordings worldwide.

---

#### How We Built the Dataset

| Step | Action |
|------|--------|
| 1 | Queried the MusicBrainz REST API across multiple genres and regions |
| 2 | Extracted artist metadata (ratings, tags, type, area) |
| 3 | Derived **Collab_Count** by counting credited artist relationships |
| 4 | Derived **Age** from career-begin dates |
| 5 | Merged with `map_data("world")` centroids for geographic coordinates |
| 6 | Cleaned and standardised country names for joining |

---

#### Variables in the Dataset

| Variable | Type | Description |
|----------|------|-------------|
| `Artist` | Categorical | Artist or group name |
| `Type` | Categorical | *Person* (solo) or *Group* |
| `Gender` | Categorical | Gender of solo artist |
| `Top_Genre` | Categorical | Primary musical genre |
| `Country` | Categorical | Country of origin |
| `Hometown` | Categorical | City/region of origin |
| `Is_Active` | Binary Categorical | Whether artist is still active |
| `User_Rating` | Numerical | Avg. user rating (0–5) |
| `Rating_Votes` | Numerical | Number of user votes |
| `Collab_Count` | Numerical | Count of credited collaborations |
| `Age` | Numerical | Years since career began |
| `Album_Count` | Numerical | Number of albums |
| `Recording_Count` | Numerical | Number of recordings |

The full code and raw data live on our
[**GitHub Repository →**](https://github.com/CelineeUWC/SDS_MP2)

Column {data-width=520}
--------------------------------------------------------------------------------

### Dataset at a Glance

```{r home-kpi}
n_artists   <- nrow(artists)
n_countries <- n_distinct(artists$Country,   na.rm = TRUE)
n_genres    <- n_distinct(artists$Top_Genre, na.rm = TRUE)
avg_rating  <- round(mean(artists$User_Rating, na.rm = TRUE), 2)

tibble(
  Metric = c("Total Artists", "Countries Represented",
             "Genres Represented", "Avg. User Rating"),
  Value  = c(
    format(n_artists, big.mark = ","),
    as.character(n_countries),
    as.character(n_genres),
    paste0(avg_rating, " / 5")
  )
) |>
  gt() |>
  tab_header(title = "Dataset Overview") |>
  cols_align("center", columns = Value) |>
  tab_style(
    style     = list(cell_fill(color = "#eef2ff"), cell_text(weight = "bold")),
    locations = cells_column_labels()
  )
```

### Sample of the Data

```{r home-table}
artists |>
  select(Artist, Country, Top_Genre, Type, Gender,
         Age, User_Rating, Rating_Votes, Collab_Count, Is_Active) |>
  slice_head(n = 15) |>
  gt() |>
  cols_label(
    Top_Genre    = "Genre",
    User_Rating  = "Rating",
    Rating_Votes = "Votes",
    Collab_Count = "Collabs",
    Is_Active    = "Active"
  ) |>
  fmt_number(columns = User_Rating, decimals = 2)
```


About
================================================================================

Column {data-width=420}
--------------------------------------------------------------------------------

### Research Questions

This project investigates **what makes a music artist popular** by analyzing
artist characteristics, geography, age, and collaboration behavior.

---

#### Popularity & Ratings
1. What factors are most associated with higher user ratings?
2. How do genre, type (solo vs group), and country relate to ratings?

####  Age & Career
3. At what age do most highly-rated **solo artists** begin their careers?
4. How are age and genre related in this data set?

####  Geography
5. Does an artist's location affect their popularity?

####  Collaboration
6. Are younger artists collaborating more than older artists?
7. How does collaboration vary by artist type (group vs person)?

Column {data-width=580}
--------------------------------------------------------------------------------

### Summary Statistics by Artist Type

```{r about-summary}
artists |>
  filter(!is.na(Type)) |>
  group_by(Type) |>
  summarize(
    Count       = n(),
    Avg_Rating  = round(mean(User_Rating,  na.rm = TRUE), 2),
    Median_Age  = median(Age,              na.rm = TRUE),
    Avg_Collabs = round(mean(Collab_Count, na.rm = TRUE), 1),
    Avg_Albums  = round(mean(Album_Count,  na.rm = TRUE), 1),
    .groups     = "drop"
  ) |>
  gt() |>
  cols_label(
    Avg_Rating  = "Avg Rating",
    Median_Age  = "Median Age",
    Avg_Collabs = "Avg Collabs",
    Avg_Albums  = "Avg Albums"
  )
```

### Genre Distribution

```{r about-genre-bar}
# EXAMPLE: horizontal bar chart coloured by count using the Viridis palette.
# reorder() makes the longest bars appear at the top.
artists |>
  count(Top_Genre, sort = TRUE) |>
  filter(!is.na(Top_Genre)) |>
  slice_head(n = 15) |>
  plot_ly(
    x           = ~n,
    y           = ~reorder(Top_Genre, n),
    type        = "bar",
    orientation = "h",
    marker      = list(color = ~n, colorscale = "Viridis", showscale = FALSE)
  ) |>
  layout(
    title  = "Top 15 Genres by Artist Count",
    xaxis  = list(title = "Number of Artists"),
    yaxis  = list(title = ""),
    margin = list(l = 150)
  )
```

### Rating by Genre — Boxplot

```{r about-genre-box}
# EXAMPLE: boxplot showing the spread of User_Rating within each genre.
# This reveals which genres have consistently high vs. variable ratings.
artists |>
  filter(!is.na(Top_Genre), !is.na(User_Rating)) |>
  mutate(Top_Genre = fct_reorder(Top_Genre, User_Rating, median)) |>
  plot_ly(
    x    = ~User_Rating,
    y    = ~Top_Genre,
    type = "box",
    orientation = "h",
    marker = list(color = "#4361ee"),
    line   = list(color = "#4361ee")
  ) |>
  layout(
    title  = "User Rating Distribution by Genre",
    xaxis  = list(title = "User Rating (0–5)", range = c(0, 5)),
    yaxis  = list(title = ""),
    margin = list(l = 150)
  )
```


Maps
================================================================================

Column {.sidebar data-width=260}
--------------------------------------------------------------------------------

**Map Filters**

```{r map-sidebar}
selectInput("map_genre", "Genre:",
            choices = all_genres, selected = "All")

selectInput("map_type", "Artist Type:",
            choices = all_types, selected = "All")

selectInput("map_active", "Active Status:",
            choices = c("All", "TRUE", "FALSE"), selected = "All")

sliderInput("map_rating", "Min. User Rating:",
            min = 0, max = 5, value = 0, step = 0.25)

hr()
p(em("Circle size = rating. Click any marker for full details."))
```

Column {data-width=700}
--------------------------------------------------------------------------------

### Interactive World Map of Artists

```{r map-leaflet}
# reactive filter
# every time a sidebar input changes, R re-runs this reactive block and the
# map below automatically redraws with only the matching rows.
map_data_r <- reactive({
  df <- artists_map |> filter(!is.na(Longitude), !is.na(Latitude))
  if (input$map_genre  != "All") df <- df |> filter(Top_Genre == input$map_genre)
  if (input$map_type   != "All") df <- df |> filter(Type      == input$map_type)
  if (input$map_active != "All")
    df <- df |> filter(as.character(Is_Active) == input$map_active)
  df <- df |> filter(is.na(User_Rating) | User_Rating >= input$map_rating)
  df
})

renderLeaflet({
  df  <- map_data_r()
  # colorNumeric maps a continuous variable (User_Rating 0–5) to a colour ramp.
  pal <- colorNumeric("YlOrRd", domain = c(0, 5), na.color = "#aaaaaa")

  leaflet(df) |>
    addProviderTiles(providers$CartoDB.Positron) |>  # clean greyscale base map
    setView(lng = 10, lat = 25, zoom = 2) |>
    addCircleMarkers(
      # jitter slightly so artists in the same country don't all stack exactly
      lng         = ~jitter(Longitude, factor = 0.8),
      lat         = ~jitter(Latitude,  factor = 0.8),
      # larger circle = higher rating
      radius      = ~ifelse(is.na(User_Rating), 5, 3 + User_Rating * 2),
      fillColor   = ~pal(User_Rating),
      fillOpacity = 0.85,
      stroke      = TRUE,
      weight      = 0.8,
      color       = "white",
      # html popup shown when user clicks a marker
      popup = ~paste0(
        "<b style='font-size:14px'>", Artist, "</b><br>",
        "<hr style='margin:4px 0'>",
        " <b>Country:</b> ",  str_to_title(Country), "<br>",
        "️ <b>Hometown:</b> ", ifelse(is.na(Hometown),  "Unknown", Hometown),  "<br>",
        " <b>Genre:</b> ",    ifelse(is.na(Top_Genre), "Unknown", Top_Genre), "<br>",
        " <b>Type:</b> ",     ifelse(is.na(Type),      "Unknown", Type),      "<br>",
        " <b>Rating:</b> ",   ifelse(is.na(User_Rating), "N/A",
                                       paste0(round(User_Rating, 2), " / 5")),  "<br>",
        " <b>Age:</b> ",      ifelse(is.na(Age), "N/A", Age),                 "<br>",
        " <b>Collabs:</b> ",  ifelse(is.na(Collab_Count), "N/A", Collab_Count)
      ),
      label = ~Artist   # tooltip shown on hover (no click needed)
    ) |>
    addLegend("bottomright",
              pal     = pal,
              values  = ~User_Rating,
              title   = "User Rating",
              opacity = 0.85)
})
```

Column {data-width=300}
--------------------------------------------------------------------------------

### Artists Shown

```{r map-vbox}
renderValueBox({
  valueBox(nrow(map_data_r()), "Artists on Map",
           icon = "fa-map-marker", color = "primary")
})
```

### Top 10 Countries (filtered)

```{r map-country-bar}
# EXAMPLE: reactive bar chart — updates whenever map filters change.
renderPlotly({
  map_data_r() |>
    count(Country, sort = TRUE) |>
    slice_head(n = 10) |>
    mutate(Country = str_to_title(Country)) |>
    plot_ly(
      x           = ~n,
      y           = ~reorder(Country, n),
      type        = "bar",
      orientation = "h",
      marker      = list(color = "#4361ee")
    ) |>
    layout(
      xaxis  = list(title = "Artist Count"),
      yaxis  = list(title = ""),
      margin = list(l = 110)
    )
})
```

### Avg. Rating by Country (top 10)

```{r map-rating-country}
# EXAMPLE: compare average rating across the most-represented countries.
renderPlotly({
  map_data_r() |>
    filter(!is.na(User_Rating)) |>
    group_by(Country) |>
    summarize(avg_rating = mean(User_Rating, na.rm = TRUE),
              n          = n(), .groups = "drop") |>
    filter(n >= 3) |>                        # only countries with 3+ artists
    arrange(desc(avg_rating)) |>
    slice_head(n = 10) |>
    mutate(Country = str_to_title(Country)) |>
    plot_ly(
      x           = ~avg_rating,
      y           = ~reorder(Country, avg_rating),
      type        = "bar",
      orientation = "h",
      marker      = list(color = ~avg_rating,
                         colorscale = "YlOrRd", showscale = FALSE)
    ) |>
    layout(
      xaxis  = list(title = "Avg. User Rating", range = c(0, 5)),
      yaxis  = list(title = ""),
      margin = list(l = 110)
    )
})
```


Artist Analysis
================================================================================

Column {.sidebar data-width=260}
--------------------------------------------------------------------------------

**Analysis Filters**

```{r analysis-sidebar}
selectInput("an_country", "Country:",
            choices = all_countries, selected = "All")

selectInput("an_type", "Artist Type:",
            choices = all_types, selected = "All")

selectInput("an_gender", "Gender:",
            choices = all_genders, selected = "All")

selectInput("an_genre", "Genre:",
            choices = all_genres, selected = "All")

sliderInput("an_votes", "Min. Rating Votes:",
            min   = 0,
            max   = max(artists$Rating_Votes, na.rm = TRUE),
            value = 0,
            step  = 5)
```

Column {data-width=500}
--------------------------------------------------------------------------------

### User Rating Distribution

```{r an-hist}
# Reactive dataset: re-filters whenever any sidebar input changes
an_data <- reactive({
  df <- artists
  if (input$an_country != "All") df <- df |> filter(Country   == input$an_country)
  if (input$an_type    != "All") df <- df |> filter(Type      == input$an_type)
  if (input$an_gender  != "All") df <- df |> filter(Gender    == input$an_gender)
  if (input$an_genre   != "All") df <- df |> filter(Top_Genre == input$an_genre)
  df <- df |> filter(is.na(Rating_Votes) | Rating_Votes >= input$an_votes)
  df
})

# EXAMPLE: histogram of User_Rating — shows whether most artists cluster
# toward the middle or at the extremes.
renderPlotly({
  plot_ly(an_data(), x = ~User_Rating, type = "histogram", nbinsx = 20,
          marker = list(color = "#4361ee",
                        line  = list(color = "white", width = 0.8))) |>
    layout(
      xaxis = list(title = "User Rating (out of 5)", range = c(0, 5)),
      yaxis = list(title = "Count")
    )
})
```


###  Album Count vs. User Rating

```{r an-albums-scatter}
# EXAMPLE: does releasing more albums correlate with a higher rating?
# Log-scale x-axis handles the right-skewed album counts better.
renderPlotly({
  df <- an_data() |> filter(!is.na(Album_Count), !is.na(User_Rating), Album_Count > 0)
  plot_ly(df,
          x     = ~Album_Count,
          y     = ~User_Rating,
          type  = "scatter",
          mode  = "markers",
          color = ~Type,
          text  = ~Artist,
          hovertemplate = "<b>%{text}</b><br>Albums: %{x}<br>Rating: %{y:.2f}<extra></extra>",
          marker = list(size = 7, opacity = 0.65)) |>
    layout(
      xaxis  = list(title = "Album Count", type = "log"),
      yaxis  = list(title = "User Rating (0–5)"),
      legend = list(title = list(text = "Type"))
    )
})
```

Column {data-width=500}
--------------------------------------------------------------------------------

###  Top 15 Artists by Rating

```{r an-top-artists}
# EXAMPLE: horizontal bar chart of the highest-rated artists.
# Only artists with at least 1 Rating_Vote are included so that an artist
# with a single 5-star vote doesn't dominate.
renderPlotly({
  an_data() |>
    filter(!is.na(User_Rating), !is.na(Artist), !is.na(Rating_Votes)) |>
    arrange(desc(User_Rating)) |>
    slice_head(n = 15) |>
    mutate(label = paste0(Artist, " (", str_to_title(Country), ")")) |>
    plot_ly(
      x           = ~User_Rating,
      y           = ~reorder(label, User_Rating),
      type        = "bar",
      orientation = "h",
      marker      = list(color      = ~User_Rating,
                         colorscale = "YlOrRd",
                         showscale  = FALSE),
      hovertemplate = "<b>%{y}</b><br>Rating: %{x:.2f}<extra></extra>"
    ) |>
    layout(
      xaxis  = list(title = "User Rating", range = c(0, 5)),
      yaxis  = list(title = ""),
      margin = list(l = 230)
    )
})
```




Collabs
================================================================================

Column {.sidebar data-width=260}
--------------------------------------------------------------------------------

**Collaboration Filters**

```{r collab-sidebar}
sliderInput("col_age", "Artist Age Range:",
            min   = floor(age_range[1]),
            max   = ceiling(age_range[2]),
            value = c(floor(age_range[1]), ceiling(age_range[2])),
            step  = 1)

selectInput("col_type", "Artist Type:",
            choices = all_types, selected = "All")

selectInput("col_genre", "Genre:",
            choices = all_genres, selected = "All")

selectInput("col_active", "Active Status:",
            choices = c("All", "TRUE", "FALSE"), selected = "All")
```

Column {data-width=580}
--------------------------------------------------------------------------------

###  Age vs. Collaboration Count

```{r collab-scatter}
col_data <- reactive({
  df <- artists |>
    filter(!is.na(Age), !is.na(Collab_Count)) |>
    filter(Age >= input$col_age[1], Age <= input$col_age[2])
  if (input$col_type   != "All") df <- df |> filter(Type      == input$col_type)
  if (input$col_genre  != "All") df <- df |> filter(Top_Genre == input$col_genre)
  if (input$col_active != "All")
    df <- df |> filter(as.character(Is_Active) == input$col_active)
  df
})

# scatterplot with an optional linear trend line.
# The trend line is built with lm() and added via add_lines().
# We guard with nrow >= 5 so the model doesn't crash on tiny subsets.
renderPlotly({
  df <- col_data()

  p <- plot_ly(df,
               x     = ~Age,
               y     = ~Collab_Count,
               type  = "scatter",
               mode  = "markers",
               color = ~Type,
               text  = ~Artist,
               hovertemplate = "<b>%{text}</b><br>Age: %{x}<br>Collabs: %{y}<extra></extra>",
               marker = list(size = 8, opacity = 0.7))

  if (nrow(df) >= 5 && sd(df$Age, na.rm = TRUE) > 0) {
    fit <- lm(Collab_Count ~ Age, data = df)
    p <- p |>
      add_lines(x          = ~Age,
                y          = fitted(fit),
                name       = "Linear trend",
                line       = list(color = "red", dash = "dash", width = 2),
                inherit    = FALSE,
                showlegend = TRUE)
  }

  p |> layout(
    xaxis  = list(title = "Artist Age (years)"),
    yaxis  = list(title = "Collaboration Count"),
    legend = list(title = list(text = "Type"))
  )
})
```



###  Collab Count vs. User Rating

```{r collab-rating-scatter}
# do artists who collaborate more also get rated higher?
# This directly addresses research question 6.
renderPlotly({
  col_data() |>
    filter(!is.na(User_Rating)) |>
    plot_ly(x     = ~Collab_Count,
            y     = ~User_Rating,
            type  = "scatter",
            mode  = "markers",
            color = ~Type,
            text  = ~Artist,
            hovertemplate = "<b>%{text}</b><br>Collabs: %{x}<br>Rating: %{y:.2f}<extra></extra>",
            marker = list(size = 7, opacity = 0.65)) |>
    layout(
      xaxis  = list(title = "Collaboration Count"),
      yaxis  = list(title = "User Rating (0–5)"),
      legend = list(title = list(text = "Type"))
    )
})
```

Column {data-width=420}
--------------------------------------------------------------------------------

###  Collab Count by Artist Type

```{r collab-box}
# EXAMPLE: boxplot comparing collab distributions for Person vs Group.
# "outliers" shows individual dots for extreme values.
renderPlotly({
  col_data() |>
    filter(!is.na(Type)) |>
    plot_ly(y         = ~Collab_Count,
            color     = ~Type,
            type      = "box",
            colors    = c("#4361ee", "#f72585"),
            boxpoints = "outliers") |>
    layout(
      yaxis  = list(title = "Collaboration Count"),
      xaxis  = list(title = "Artist Type"),
      legend = list(title = list(text = "Type"))
    )
})
```


### Summary Table

```{r collab-table}
renderTable({
  col_data() |>
    filter(!is.na(Type)) |>
    group_by(Type) |>
    summarize(
      Artists       = n(),
      `Avg Collabs` = round(mean(Collab_Count, na.rm = TRUE), 2),
      Median        = median(Collab_Count,     na.rm = TRUE),
      Max           = max(Collab_Count,        na.rm = TRUE),
      .groups       = "drop"
    )
}, striped = TRUE, hover = TRUE, bordered = TRUE)
```


Conclusion
================================================================================

Column {data-width=500}
--------------------------------------------------------------------------------

###  Key Findings

---

#### Geographic Distribution
- Artists are concentrated in the **USA**, **UK**, and **Canada**, reflecting
  MusicBrainz's stronger coverage of Western artists.
- Country of origin shows a weak direct relationship with rating;
  **genre** is a stronger predictor of popularity.

---

#### Popularity & Ratings
- **Genre** is among the strongest predictors of user ratings — certain genres
  consistently attract higher scores.
- Artists with few `Rating_Votes` display extreme highs and lows; those with
  many votes converge toward the mid-range (regression to the mean).
- No strong relationship was found between **age** and rating alone.

---

####  Age & Career
- Most highly-rated solo artists began their careers in their **20s**.
- Older artists show slightly higher collaboration counts,
  consistent with longer time in the industry.

---

####  Collaboration Patterns
- **Groups** tend to have higher `Collab_Count` than solo artists.
- Collaboration count shows only a weak relationship with user rating,
  suggesting popularity is driven more by genre and reach than volume of work.

Column {data-width=500}
--------------------------------------------------------------------------------

### ️ Limitations

| Limitation | Detail |
|------------|--------|
| Western bias | MusicBrainz has denser coverage of US/UK artists |
| Crowd-sourced ratings | Ratings reflect listener demographics, not objective quality |
| Our favorite artists| We picked the artists in this dataset |
| Collab undercounting | Informal collaborations not in MusicBrainz are excluded |
| Country cleaning | We were only allowed 25 requests and therefore there is some counts limited to 25|



### References

- [MusicBrainz API](https://musicbrainz.org/doc/MusicBrainz_API)
- [flexdashboard](https://pkgs.rstudio.com/flexdashboard/)
- [Leaflet for R](https://rstudio.github.io/leaflet/)
- [plotly for R](https://plotly.com/r/)
- [GitHub Repository](https://github.com/CelineeUWC/SDS_MP2)
